---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# seqUtils

 <!-- badges: start -->
  [![R-CMD-check](https://github.com/SamT123/seqUtils/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/SamT123/seqUtils/actions/workflows/R-CMD-check.yaml)
  <!-- badges: end -->

Utilities for working with biological sequences in R.

## Overview

`seqUtils` provides simple, efficient tools for common sequence analysis tasks:

- **Read/write FASTA files** - Fast I/O for single-line and multi-line FASTA formats
- **Translate sequences** - DNA/RNA to amino acids with deletion handling
- **Align sequences** - Wrapper for MAFFT alignment to reference sequences
- **Find substitutions** - Compare sequences and identify differences
- **Calculate consensus** - Generate consensus sequences from alignments
- **Clean sequences** - Remove or replace non-standard characters

## Installation

You can install the development version of seqUtils from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("SamT123/seqUtils")
```

## Example

```{r example, eval=FALSE}
library(seqUtils)

# Read FASTA file
sequences <- fast_fasta("sequences.fasta")

# Translate to amino acids
aa_sequences <- translate(sequences)

# Find substitutions compared to a reference
substitutions <- get_substitutions(
  reference_seq,
  query_seqs,
  exclude = c("X", "N")
)
```

## Requirements

- Biostrings package
- MAFFT (for alignment functions)
